---
title: "HydroCO2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(tidyverse, httr, lubridate, janitor, jsonlite, listviewer, usethis, rvest, data.table, xml2, readxl, here, dplyr, plotly, lubridate, naniar, readr, broom)
```

### Data Wrangling

## Importing the data, summarizing
#*note* the data was imported using the "import dataset" tab in Rstudio. 

```{r hydro}
hydro_data <- readr::read_csv("/Users/EthanMcChristian/Desktop/hydro_co2.csv")


```

## Cleaning column names


```{r names, echo=TRUE}
hydro_data <- clean_names(hydro_data)
hydro_data
```
```{r}
hydro_data <- 
  hydro_data %>%
  rename(
 co2_percap = co2_emissions_metric_tons_per_capita_en_atm_co2e_pc,
 perc_elec = access_to_electricity_percent_of_population_eg_elc_accs_zs,
 gni_percap = gni_per_capita_atlas_method_current_us_ny_gnp_pcap_cd,
 co2_emission = co2_emissions_kt_en_atm_co2e_kt,
 perc_co2_gaseous = co2_emissions_from_gaseous_fuel_consumption_percent_of_total_en_atm_co2e_gf_zs,
 perc_co2_solid_fuel = co2_emissions_from_solid_fuel_consumption_percent_of_total_en_atm_co2e_sf_zs,
 perc_co2_liquid_fuel = co2_emissions_from_liquid_fuel_consumption_percent_of_total_en_atm_co2e_lf_zs,
 co2_methane_equiv.energy_sector = methane_emissions_in_energy_sector_thousand_metric_tons_of_co2_equivalent_en_atm_meth_eg_kt_ce,
 co2_methane_equiv.agriculture = agricultural_methane_emissions_thousand_metric_tons_of_co2_equivalent_en_atm_meth_ag_kt_ce,
 co2_total_greenhouse_equiv = total_greenhouse_gas_emissions_kt_of_co2_equivalent_en_atm_ghgt_kt_ce,
 perc_safe_water = people_using_safely_managed_drinking_water_services_percent_of_population_sh_h2o_smdw_zs,
 hydro_electric_perc = electricity_production_from_hydroelectric_sources_percent_of_total_eg_elc_hyro_zs,
 coal_electric_perc = electricity_production_from_coal_sources_percent_of_total_eg_elc_coal_zs,
 natural_gas_electric_perc = electricity_production_from_natural_gas_sources_percent_of_total_eg_elc_ngas_zs,
 nuclear_electric_perc = electricity_production_from_nuclear_sources_percent_of_total_eg_elc_nucl_zs,
 oil_gas_coal_electric_perc = electricity_production_from_oil_gas_and_coal_sources_percent_of_total_eg_elc_fosl_zs,
 renewable_electric_nohydro_perc = electricity_production_from_renewable_sources_excluding_hydroelectric_percent_of_total_eg_elc_rnwx_zs,
 electric_consumption_percap = electric_power_consumption_k_wh_per_capita_eg_use_elec_kh_pc,
 pop = population_total_sp_pop_totl
  )
hydro_data
```

# Fixing my name mistakes
```{r}
hydro_co2 <- hydro_co2 %>%
  rename(
    perc_safe_water = hydro_electric_perc,
    hydro_electric_perc = electricity_production_from_hydroelectric_sources_of_total_eg_elc_hyro_zs
  )
```

## Deleting all rows with NA files in a new dataframe
# Replacing all ".." with NA's
```{r}
hydro_data <- hydro_data %>%
  mutate(co2_percap = na_if(co2_percap, "..")) %>%
  mutate(perc_elec = na_if(perc_elec, "..")) %>%
  mutate(gni_percap = na_if(gni_percap, "..")) %>%
  mutate(co2_emission = na_if(co2_emission, "..")) %>%
  mutate(perc_co2_gaseous = na_if(perc_co2_gaseous, "..")) %>%
  mutate(perc_co2_solid_fuel = na_if(perc_co2_solid_fuel, "..")) %>%
  mutate(perc_co2_liquid_fuel = na_if(perc_co2_liquid_fuel, "..")) %>%
  mutate(co2_methane_equiv.energy_sector = na_if(co2_methane_equiv.energy_sector, "..")) %>%
  mutate(co2_methane_equiv.agriculture = na_if(co2_methane_equiv.agriculture, "..")) %>%
  mutate(co2_total_greenhouse_equiv = na_if(co2_total_greenhouse_equiv, "..")) %>%
  mutate(perc_safe_water = na_if(perc_safe_water, "..")) %>%
  mutate(hydro_electric_perc = na_if(hydro_electric_perc, "..")) %>%
  mutate(coal_electric_perc = na_if(coal_electric_perc, "..")) %>%
  mutate(natural_gas_electric_perc = na_if(natural_gas_electric_perc, "..")) %>%
  mutate(nuclear_electric_perc = na_if(nuclear_electric_perc, "..")) %>%
  mutate(oil_gas_coal_electric_perc = na_if(oil_gas_coal_electric_perc, "..")) %>%
  mutate(renewable_electric_nohydro_perc = na_if(renewable_electric_nohydro_perc, "..")) %>%
  mutate(electric_consumption_percap = na_if(electric_consumption_percap, "..")) %>%
  mutate(pop = na_if(pop, ".."))

```
# Deleting all rows with NA's, converting all factor variables to numeric
```{r}
hydro_data_subset <- hydro_data[complete.cases(hydro_data), ]
hydro_data_subset
```
```{r}
hydro_data_subnum <- hydro_data_subset %>%
  mutate(co2_percap = as.numeric(co2_percap)) %>%
  mutate(perc_elec = as.numeric(perc_elec)) %>%
  mutate(gni_percap = as.numeric(gni_percap)) %>%
  mutate(co2_emission = as.numeric(co2_emission)) %>%
  mutate(perc_co2_gaseous = as.numeric(perc_co2_gaseous)) %>%
  mutate(perc_co2_solid_fuel = as.numeric(perc_co2_solid_fuel)) %>%
  mutate(perc_co2_liquid_fuel = as.numeric(perc_co2_liquid_fuel)) %>%
  mutate(co2_methane_equiv.energy_sector = as.numeric(co2_methane_equiv.energy_sector)) %>%
  mutate(co2_methane_equiv.agriculture = as.numeric(co2_methane_equiv.agriculture)) %>%
  mutate(co2_total_greenhouse_equiv = as.numeric(co2_total_greenhouse_equiv)) %>%
  mutate(perc_safe_water = as.numeric(perc_safe_water)) %>%
  mutate(hydro_electric_perc = as.numeric(hydro_electric_perc)) %>%
  mutate(coal_electric_perc = as.numeric(coal_electric_perc)) %>%
  mutate(natural_gas_electric_perc = as.numeric(natural_gas_electric_perc)) %>%
  mutate(nuclear_electric_perc = as.numeric(nuclear_electric_perc)) %>%
  mutate(oil_gas_coal_electric_perc = as.numeric(oil_gas_coal_electric_perc)) %>%
  mutate(renewable_electric_nohydro_perc = as.numeric(renewable_electric_nohydro_perc)) %>%
  mutate(electric_consumption_percap = as.numeric(electric_consumption_percap)) %>%
  mutate(pop = as.numeric(pop))
```
NOTE: the data set has shrunk from 3965 observations to only 385, which means we deleted a LOT there. With this newer model, we have certaintly introduced selection (or reporting) bias. This means that countries that did not report any number of inputs is not counted in the regression. Are the countries who did not report a value more likely to have illecit greenhouse gas emissions taking place?

## time to convert my "time" values to time series data
```{r}
hydro_data_subnum <- hydro_data_subnum %>%
  mutate(time = as.character(time))
hydro_data_subnum_date <- hydro_data_subnum %>%
  mutate(time = as.Date(time, "%Y"))
```
## Running a fun little regression, just to see
```{r}
reg1 <- lm(log(co2_total_greenhouse_equiv) ~ hydro_electric_perc + log(pop), data = hydro_data_subnum_date)
tidy(reg1)
```
## Adding a "percapital greenhouse gas emissions in kt of co2 equivalent" column. 
```{r}
hydro_data_new <- hydro_data_subnum_date %>%
  mutate(percap_greenhouse = co2_total_greenhouse_equiv/pop)
hydro_data_new
```
## Filtering data by middle income countries (UN World Economic Situation and Prospects) and by countries with access to drinking water equal to or above the world average. 
```{r}
hydro_data_new2 <-hydro_data_new %>%
  filter(gni_percap >= 996) %>%
  filter(perc_safe_water >= (530/7.9))
hydro_data_choice <- hydro_data_new2 %>%
  filter(country_name %in% c("Central Europe and the Baltics", "Euro Area", "Europe and Central Asia", "Europe and Central Asia (excluding high income)", "Europe and Central Asia (IDA and IBRD countries)",  "European Union", "High income", "North America", "OECD members", "Post-demographic dividend"))

```
```{r}
reg2 <- lm(percap_greenhouse ~ hydro_electric_perc + coal_electric_perc + nuclear_electric_perc + natural_gas_electric_perc + oil_gas_coal_electric_perc + renewable_electric_nohydro_perc + log(electric_consumption_percap) + log(pop), data = hydro_data_new2)
tidy(reg2)

ggplot(data = hydro_data_new2, aes(x = hydro_electric_perc, y = co2_total_greenhouse_equiv)) + 
       geom_line() + 
       geom_point(alpha = .5)
```